%%%%%%% Lecture Notes Style for Concrete Mathematics %%%%%%%%%%%%%
%%%%%%% Taught by Patrick White, TJHSST %%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,twosided]{article}
\input{concrete-header.tex}

%%%% add package inclusions here, if any

%%%%% Define your custom functions and macros here, if any

%%%%%%%%%%%%%%%%%%%% Define the variables below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\titlestring{The Lecture Title}
\def\scribestring{Your Name}
\def\datestring{Day, Mon, Date Year}


%%%%%%%%%%%%%%%%%%% Page Headers -- Do Not Change %%%%%%%%%%%%%%%%%%%%%%%%%%%
\lhead{\titlestring}
\rhead{Page \thepage}
\cfoot{Concrete Math -- White -- TJHSST}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

%%%%%%%%%%%%%%%%% Begin the document %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{plain}  %% no headers on this page

%%%% Do not change these lines %%%%%
\noindent
{\LARGE \textbf{\titlestring}}\\\\
%
{\Large Scribe: \scribestring}\\ \\
{\textbf{Date}: \datestring}


%%%%%%%%%%%%%%%%%%%%%%%%%% YOUR CONTENT GOES HERE %%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent

\section{Triangulation}
\begin{problem}{1}
How many triangulations are there of a labeled hexagon? 
\end{problem}
Note for a square, there are two  (one that is just rotated):  
% insert diagram
And for a pentagon, there are five - just one possible configuration that can be rotated: 
% insert diagram
\begin{proof}[Solution]
% insert diagrams
This gives a total of $\boxed{14}$ total triangulations. 
\end{proof}
\begin{proof}
We can solve the more general case first. Let $T_n$ be the number of ways to triangulate an $n$-gon. Notice that if we draw a diagonal of the $n$-gon 

\end{proof}

\section{Laser Grid}
\begin{problem}
	Consider walking on the coordinate plane, but you are constrained to either walk over by one 
\end{problem}

\begin{solution}

\end{solution}

\begin{problem}{2}
	Consider walking on the coordinate plane, but you are constrained to either walk 
\end{problem}

\begin{solution}
We will employ complementary counting by counting how many paths cross the line. The total number of paths (from the first problem) is known to be $\binom{2n}{n}$. 

Now, define a \textit{bad path} to be one that crosses the axis. Note that any bad path must touch the line $y = -1$. We will reflect the part of the path from $(0, 0)$ to the first intersection with $y = -1$ over the line $y = -1$. Doing so yields a path from $(0, -2)$ to $(2n, 0)$ with no restrictions. 

We now claim that all bad paths from $(0, -2)$, when flipped in this manner, will yield a bad path in the original problem. This path must cross the line $y = -1$ at least once. Notice that when we undo this reflection, the path will now start at $(0, 0)$, and is forced to touch the line $y = -1$, and this uniquely determines exactly one bad path. Therefore, it suffices to count the number of paths from $(0, -2)$ to 

Finally, the number of good paths is $\binom{2n}{n} - \binom{2n}{n-1}$. This can be simplified to yield $\frac{1}{n+1} \binom{2n}{n}$, which is the $n$th Catalan number. 
\end{solution}

Remember also that we encountered these before in the binomial expansion of $(1-x)^\frac{1}{2}$! 



\end{document}
