%%%%%%% Lecture Notes Style for Concrete Mathematics %%%%%%%%%%%%%
%%%%%%% Taught by Patrick White, TJHSST %%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,twosided]{article}
\input{concrete-header.tex}

%%%% add package inclusions here, if any
\usepackage{blubase}
%%%%% Define your custom functions and macros here, if any

%%%%%%%%%%%%%%%%%%%% Define the variables below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\titlestring{The Lecture Title}
\def\scribestring{Your Name}
\def\datestring{Day, Mon, Date Year}


%%%%%%%%%%%%%%%%%%% Page Headers -- Do Not Change %%%%%%%%%%%%%%%%%%%%%%%%%%%
\lhead{\titlestring}
\rhead{Page \thepage}
\cfoot{Concrete Math -- White -- TJHSST}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

%%%%%%%%%%%%%%%%% Begin the document %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{plain}  %% no headers on this page

%%%% Do not change these lines %%%%%
\noindent
{\LARGE \textbf{\titlestring}}\\\\
%
{\Large Scribe: \scribestring}\\ \\
{\textbf{Date}: \datestring}


%%%%%%%%%%%%%%%%%%%%%%%%%% YOUR CONTENT GOES HERE %%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent

\section{Catalan's Reprise}
Recall we wanted to count the number of triangulations of a regular polygon with $n$ sides - let $T_n$ be the number of ways to triangulate a $n$-gon. 

The recursive process is constructed by fixing an edge that must be in exactly one triangle, and then doing casework on which of the other $n-2$ vertices it must have. Suppose we label the vertices of the $n$-gon from 1 to $n$ and have our fixed edge be the edge with vertices labeled $1$ and $n$. If we pick vertex $k$ to be the other vertex, we can triangulate the two other polygons, one in $T_k$ and another in $T_{n-k-1}$ ways. Therefore, we have the recurrence relation
\[
	T_n = \sum_{k=2}^{n-1} T_k T_{n-k+1}
\]

Shifting the indices back by 2, it appears to be that $C_n$ is the number of ways to triangulate an $n+2$-gon. To rigorously show this, we will have to manipulate the recurrence relation - and by doing this we see that 
\[
	C_{n+1} = \sum_{k=1}^n C_k C_{n-k}
\]
Fundamentally, we will see the Catalan numbers reappear when we can break up a problem into two smaller subproblems. This is not extremely obvious for the path problem, but we can consider the first point at which the path crosses the line $y=1$ again?

\section{Equivalence Relation Problems}
An equivalence relation $\cong$ if it is symmetric, reflexive, and transitive. A general relation on a set $A$ is a subset of the set $A \times A$. We can describe this relation as a table, called a \textbf{Cayley Table}. Here is a relation on 4 elements: 

%x a b c d
%a     x  
%b   x x
%c 
%d       x

As another example, the "greater than" and "less than" relations are relations that 

\begin{problem}
How many relations are there on a set of $n$ elements? How many relations are reflexive? How many are reflexive and symmetric? How many are reflexive, symmetric, and transitive? 
\end{problem}

\begin{solution}
	There are clearly $2^{n^2}$ possible relations. 
	
	For a reflexive relation, we just fix the main diagonal. This gives $2^{n^2-n}$. 
	
	For reflexive and symmetric relations, we not only must fix the main diagonal, but then out of any of the $\binom{n}{2}$ 
	
	For reflexive, symmetric, and transitive relations, we must think of this differently. 
	
	The answer is $\sum_{k=1}^n \stirii{n}{k} = B_n$, the $n$th \textbf{Bell  number}. 
\end{solution}
\end{document}
