\label{20-0319-2}
% %%%%%%% Lecture Notes Style for Concrete Mathematics %%%%%%%%%%%%%
% %%%%%%% Taught by Patrick White, TJHSST %%%%%%%%%%%%%%%%%%%%%%%%%%

% \documentclass[11pt,twosided]{article}
% \input{concrete-header.tex}

% %%%% add package inclusions here, if any
% \usepackage{ textcomp }
% \usepackage{amsmath}
% %%%%% Define your custom functions and macros here, if any

% %%%%%%%%%%%%%%%%%%%% Define the variables below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \def\titlestring{Generating Functions \& Solving Recurrences}
% \def\scribestring{Ritika Shrivastav, Anupama Jayaraman}
% \def\datestring{19, March, 2019}


% %%%%%%%%%%%%%%%%%%% Page Headers -- Do Not Change %%%%%%%%%%%%%%%%%%%%%%%%%%%
% \lhead{\titlestring}
% \rhead{Page \thepage}
% \cfoot{Concrete Math -- White -- TJHSST}
% \renewcommand{\headrulewidth}{0.4pt}
% \renewcommand{\footrulewidth}{0.4pt}
% \newcommand{\}{\genfrac{\lbrace}{\rbrace}{0pt}{}}
% %%%%%%%%%%%%%%%%% Begin the document %%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{document}
% \thispagestyle{plain}  %% no headers on this page

% %%%% Do not change these lines %%%%%
% \noindent
% {\LARGE \textbf{\titlestring}}\\\\
% %
% {\Large Scribe: \scribestring}\\ \\
% {\textbf{Date}: \datestring}


%%%%%%%%%%%%%%%%%%%%%%%%%% YOUR CONTENT GOES HERE %%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent

\subsection{More Applications of Generating Functions}

\subsection*{A food counting problem}
\begin{problem}1 Fill a bag with $n$ pieces of fruit \underline{GIVEN}: The number of apples is even, the number of bananas is a multiple of 5, the number of oranges is at most 4, and the number of pears is at most 1. How many ways can I do this? \end{problem}
\begin{solution}
\\To solve this problem, let's find find a generating function that represents the situation. To do so, let's breakdown the function into its apples, bananas, oranges, and pears terms:\\ \\
Apples: We want 0, 2, 4, ... apples. We can write this as $(x^0 + x^2 + x^4 + x^6 +...) = (1 + x^2 + x^4 + x^6 +...)$. This is also equal to $\frac{1}{1-x^2}$.\\\\
Bananas: We want 0, 5, 10, ... oranges. We can write this as $(1 + x^5 + x^10 + x^15 +...)$, or $\frac{1}{1-x^5}$.\\\\
Oranges: We want 0, 1, 2, 3, or 4 oranges. We can write this as $(1 + x + x^2 + x^3 + x^4)$. To write this as a fraction, let's make use of the formula for a truncated series: $(1-x^{n+1}) = (1-x)(1+x+x^2+...+x^n)$. From this formula, we can see that we can write the number oranges as $\frac{1-x^5}{1-x}$. \\\\
Pears: We want either 0 or 1 pear, or $(1+x)$.\\\\\
If we multiply all of the components together, we get: \\\\
$g(x) = (1 + x^2 + x^4 + x^6 +...)(1 + x^5 + x^10 + x^15 +...)(1 + x + x^2 + x^3 + x^4)(1+x) = \frac{1}{1-x^2}*\frac{1}{1-x^5}*\frac{1-x^5}{1-x}*(1+x) = \frac{1}{(1-x)^2}$ \\\\
$\frac{1}{(1-x)^2}$ has the form of the summation of a multiset: $\sum_{k = 0}^\infty \multiset{2}{k}x^k = \sum_{k = 0}^\infty (k+1)x^k$. To find the number of ways to fill the bag with fruit, we simply find the coefficient of $x^n$ in the expansion of the summation. In other words, there are:\\\\
\framebox{$(n+1)$ ways}.

\end{solution}

\subsection*{Making Change for \$1}
\begin{problem}2 Write a generating function for making change given 1\textcent{}, 5\textcent{}, 10\textcent{}, 25\textcent{}, and 50\textcent{} coins. 
\end{problem}
\begin{solution} 
\\As with the fruit problem, let's create a generating function for the scenario. We will be working in terms of cents.\\\\
1\textcent{}: $(1+x+x^2+x^3+...)=\frac{1}{1-x}$ \\\\
5\textcent{}: $(1+x^5+x^10+...)=\frac{1}{1-x^5}$\\\\
10\textcent{}: $(1+x^10+x^20+...)=\frac{1}{1-x^{10}}$\\\\
25\textcent{}: $(1+x^25+x^50+...)=\frac{1}{1-x^{25}}$\\\\
50\textcent{}: $(1+x^50+x^100+...)=\frac{1}{1-x^{50}}$\\\\
By multiplying all of the terms together, we have the generating function: \\\\
$g(x)=\frac{1}{(1-x)(1-x^5)(1-x^{10})(1-x^{25})(1-x^{50})}$ \\\\
To find the number of ways we can make change for \$1, we find the coefficient of $x^{100}$. This is \framebox{293}.

What if there are coins available in every single denomination?
Then we have... \\\\
1\textcent{}: $(1+x+x^2+x^3+...)=\frac{1}{1-x}$ \\\\
2\textcent{}: $(1+x^2+x^4+...)=\frac{1}{1-x^2}$\\\\
3\textcent{}: $(1+x^3+x^6+...)=\frac{1}{1-x^3}$\\\\
4\textcent{}: $(1+x^4+x^8+...)=\frac{1}{1-x^4}$\\\\
n\textcent{}: $(1+x^n+x^2n0+...)=\frac{1}{1-x^n}$\\\\
By multiplying all of the terms together, we have the generating function: \\\\
$g(x)=\frac{1}{(1-x)(1-x^2)(1-x^3)(1-x^4)(1-x^n)}$ \\\\
This is also the generating function for integer partitions. \\\\
$g(x) = \prod_{k=0}^n \frac{1}{1-x^k} = \sum_{r>0} P(r)x^r $ \\\\
P(r) is the is the number of integer partitions of r.
\end{solution}

%extension --> every denomination


\begin{theorem}The number of partitions of n into distinct parts $r$ is the number of partitions of $n$ into odd parts.\end{theorem}

Distinct parts is when you split the number into parts but the sum doesn't include 2 of the same number. For example, $8 = 1 + 7$ is distinct because 1 and 7 are different. However $8 = 4 + 2 + 2$ is not because you have two 2's. \\\\
Odd parts are those that are made of a repeated sum of only odd numbers, that can be repeated as many times as one pleases. For example, $8 = 3 + 3 + 1 + 1$ is fine because it contains only odds. However, $8 = 3 + 3 + 2$ is not because 2 is an even number.\\\\
Generating Functions:\\\\
Distinct parts: $g(x) = (1+x)(1+x^2)(1+x^3)...$\\\\
Odd parts: $g(x) = (\frac{1}{1-x})(\frac{1}{1-x^3})(\frac{1}{1-x^5})...$\\\\
Claim they are identical. So set both generating functions equal and use the trick that $1+x = \frac{1-x^2}{1-x}$ \\\\
$(1+x)(1+x^2)(1+x^3)... = (\frac{1}{1-x})(\frac{1}{1-x^3})(\frac{1}{1-x^5})...$\\\\
$(\frac{1-x^2}{1-x})(\frac{1-x^4}{1-x^2})(\frac{1-x^6}{1-x^3})... = (\frac{1}{1-x})(\frac{1}{1-x^3})(\frac{1}{1-x^5})...$\\\\
We notice that term cancel neatly to provide that the two sides are directly equal and therefore this theorem has been proved.

\section{Solving Recurrences }
\subsection{Intro to Solving Recurrence Equations}

\begin{problem} 3
Given $a_n = a_{n-1} + 8a_{n-2} - 12a_{n-3}$; $a_0 = 2, a_1 = 3$, $a_2 = 19$, find the generating function $g(x) = \sum_{n = 0}^\infty a_nx^n$ and formula for $a_n$.
\end{problem}
\begin{solution}

\noindent Let's start with $g(x) = \sum_{n = 0}^\infty a_nx^n$\\\\
\noindent
Since we know the first three terms from our initial condition, we can add up our base cases and the summation of $a_n$ from $n=3$ to $n=\infty$. \\

$g(x) = 2 + 3x + 19x^2 + \sum_{n = 3}^\infty a_nx^n$ \\

\noindent
We know the recurrence relation so we can substitute it into the equation we have above\\

$g(x) = 2 + 3x + 19x^2 + \sum_{n = 3}^\infty (a_{n-1} + 8a_{n-2} - 12a_{n-3})x^n$\\

\noindent
and expand it properly\\

$g(x) = 2 + 3x + 19x^2 + \sum_{n = 3}^\infty a_{n-1}x^n + \sum_{n = 3}^\infty 8a_{n-2}x^n - \sum_{n = 3}^\infty 12a_{n-3}x^n$ \\

\noindent
Now, we need to make some manipulations. First, we want the subscript of each of the $a_n$'s to match the power to which each x is being raised. To do so, let's factor out the coefficients and the appropriate number of $x$'s from each term.\\

$g(x) = 2 + 3x + 19x^2 + x\sum_{n = 3}^\infty a_{n-1}x^{n-1} + 8x^2\sum_{n = 3}^\infty a_{n-2}x^{n-2} - 12x^3\sum_{n = 3}^\infty a_{n-3}x^{n-3}$ \\

\noindent
The second manipulation involves shifting the indices such that the term within the summation is $a_nx^n$. \\

$g(x) = 2 + 3x + 19x^2 + x\sum_{n = 2}^\infty a_{n}x^{n} + 8x^2\sum_{n = 1}^\infty a_{n}x^{n} - 12x^3\sum_{n = 0}^\infty a_{n}x^{n}$ \\

\noindent
Now that we have the new summations in terms of $a_nx^n$, they can all be rewritten using the original $g(x)$ function minus the initial terms that are excluded from the summation:\\

$g(x) = 2 + 3x + 19x^2 + x(g(x) - a_0 - a_1x) + 8x^2(g(x) - a_0) - 12x^3g(x)$
\\ \indent \indent $ = 2 + 3x + 19x^2 + x(g(x) - 2 - 3x) + 8x^2(g(x) - 2) - 12x^3g(x)$ \\

\noindent
Given this equation, we can finally solve for $g(x)$ by moving terms around. After we do this, we can see that our final equation for $g(x)$ is \\

\framebox{\textbf{$g(x) = \frac{2 + x}{1 - x - 8x^2 - 12x^3}$}}\\

\noindent\normalsize
\textbf{**Notice}: \textit{If we move all of the terms in the original $a_n$ equation to one side, we obtain $a_n - a_{n-1} - 8a_{n-2} + 12a_{n-3} = 0$. This resembles the denominator of our final $g(x)$ equation.} 

\noindent
Now let's find the formula for $a_n$. With our equation for $g(x)$, we can factor the denominator into $(1 + 3x)(1 - 2x)^2$ to determine an equation for $a_n$. More specifically, we can use \textit{partial fraction decomposition} to breakdown the function into terms that will be more useful to us.\\

$g(x) = \frac{2 + x}{1 - x - 8x^2 - 12x^3} = \frac{A}{1 + 3x} + \frac{B}{1 - 2x} + \frac{C}{(1-2x)^2}$ \\

\noindent
To conduct partial fraction decomposition, we multiply the denominator of $g(x)$ on both sides of the equation.\\

$2 + x = A(1-2x)^2 + B(1 + 3x)(1 - 2x) + C(1 + 3x)$\\

\noindent
Once we solve this equation, by plugging in different values for $x$, we get 
$A = \frac{3}{5}, B = \frac{2}{5}, C = 1$. This allows as to rewrite $g(x)$ as a sum of fractions. \\

{$g(x) = \frac{3}{5}(\frac{1}{1 + 3x}) + \frac{2}{5}(\frac{1}{1 - 2x}) + \frac{1}{(1-2x)^2}$}\\

\noindent\normalsize
Looking closer, we notice that each fraction can be rewritten as a summation. We can convert the equations by using the following principles:\\

1) $\frac{1}{1 - x} = \sum_{n = 0}^\infty x^n$ and 
2) $\frac{1}{(1 - x)^k} = \sum_{n = 0}^\infty \multiset{n}{k}x^n$\\

\noindent
With these principles, we get the following equation:\\

$g(x) = \frac{3}{5}\sum_{n = 0}^\infty (-3x)^n + \frac{2}{5}\sum_{n = 0}^\infty 2x^n + \sum_{n = 0}^\infty (n + 1)(2x)^n$\\

\noindent
We can now use this equation to achieve our final equation for $a_n$. To do so, we simply remove the summations and the $x^n$'s from the equation.:\\

\framebox{\textbf{$a_n = \frac{3}{5}(-3)^n + \frac{2}{5}(2^n) + (n + 1)2^n$}} 

\end{solution} \\



\begin{problem} 4
Given $a_n = 6a_{n-1} - 9a_{n-2}$; $a_0 = 1, a_1 = 9$, find $a_n$.
\end{problem}
\begin{solution}
We find the generating function....\\\\
$g(x) = \sum_{n = 0}^\infty a_nx^n$\\\\
$g(x) = 1 + 9x + \sum_{n = 2}^\infty (6a_{n-1} - 9a_{n-2})x^n$\\\\
$g(x) = 1 + 9x + \sum_{n = 2}^\infty 6a_{n-1}x^n - \sum_{n = 2}^\infty 9a_{n-2}x^n$\\\\
$g(x) = 1 + 9x + 6x\sum_{n = 2}^\infty 6a_{n-1}x^{n-1} - 9x^2\sum_{n = 2}^\infty a_{n-2}x^{n-2}$\\\\
$g(x) = 1 + 9x + 6x\sum_{n = 1}^\infty 6a_{n}x^{n} - 9x^2\sum_{n = 0}^\infty a_{n}x^{n}$\\\\
$g(x) = 1 + 9x + 6x(g(x) - a_0) - 9x^2g(x)$\\\\\\
$g(x)(1 - 6x - 9x^2) = 1 + 9x - 6x$\\\\
$g(x) = \frac{1 + 3x}{1 - 6x - 9x^2}$\\\\
Then, we use partial fractions...\\\\
$\frac{1 + 3x}{1 - 6x - 9x^2} = \frac{A}{1 - 3x} + \frac{B}{(1 - 3x)^2}$\\\\
$1 + 3x = A(1 - 3x) + B$\\\\
If $x = 1/3$...$2 = B$\\
If $x = 0$...$1 = A + 2$, $A = -1$\\\\
This gives us an expression like so...\\\\
$g(x) = -\frac{1}{1 - 3x} + \frac{2}{(1 - 3x)^2}$\\\\
Using summation principles mentioned above, we get...\\\\
$g(x) = -1\sum_{n = 0}^\infty (3x)^n + 2\sum_{n = 0}^\infty (n + 1)(3x)^n$\\\\
$a_n = (-1)3^n + 2(n + 1)3^n$\\\\
$a_n = 3^n(-1 + 2n + 2)$\\\\
and\\\\
$a_n = (2n + 1)3^n$

\end{solution} \\


\begin{problem} 5
given $a_n = -3a_{n-1} + 10a_{n-2} + 3(2^n)$; $a_0 = 1, a_1 = 6$, find $a_n$.
\end{problem}
\begin{solution}
We will use the steps from above to find $a_n$. \\

\noindent
First let's find the generating function: \\\\
$g(x) = 0 + 6x + \sum_{n = 2}^\infty a_n x^n$ \\\\
$g(x) = 6x + \sum_{n = 2}^\infty(-3a_{n-1} + 10a_{n-2} + 3(2^n))x^n$\\\\
$g(x) = 6x + -3\sum_{n = 2}^\infty a_{n-1}x^n + 10\sum_{n = 2}^\infty a_{n-2}x^n + 3\sum_{n = 2}^\infty (2x)^n$\\\\
$g(x) =  6x - 3x\sum_{n = 2}^\infty a_{n-1}x^{n-1} + 10x^2\sum_{n = 2}^\infty a_{n-2}x^{n-2} + 3\sum_{n = 2}^\infty(2x)^n$\\\\
$g(x) =  6x - 3x\sum_{n = 1}^\infty a_{n}x^{n} + 10x^2\sum_{n = 0}^\infty a_{n}x^{n} + 3\sum_{n = 2}^\infty(2x)^n$\\\\
$g(x) =  6x - 3x(g(x) - 0) + 10x^2(g(x)) + 3(\frac{1}{1-2x}-1-2x)$\\\\
$g(x)[1+3x-10x^2]=6x+\frac{3}{1-2x}-3-6x=\frac{6x}{1-2x}$ \\\\
$g(x)=\frac{6x}{(1-2x)(1+3x-10x^2)}=\frac{6x}{(1-2x)^2(1+5x)}$ \\\\
\noindent \\
Now for $a_n$:\\\\
$\frac{6x}{(1-2x)^2(1+5x)} = \frac{A}{1-2x}+\frac{B}{(1-2x)^2}+\frac{C}{1+5x}$\\\\
\noindent
By solving this partial fraction equation, we get $A=\frac{-12}{49}$, $B=\frac{6}{7}$, and $C=\frac{-30}{49}$. Since these are messy numbers, we will hold off on substituting them into our equations until the end. \\\\
$g(x) =  A\sum_{n = 0}^\infty 2^n x^n + B\sum_{n = 0}^\infty (n+1)2^n x^{n} + C\sum_{n = 0}^\infty(-5)^n x^n$\\\\
$a_n = A*2^n + B(n+1)2^n + C(-5)^n$ \\\\
\noindent
Once we plug in the values of A, B, and C into the equation, we get our final equation for $a_n$: \\\\
\framebox{$a_n = \frac{-12}{49}2^n + \frac{6}{7}(n+1)2^n + \frac{-30}{49}(-5)^n$}

\end{solution} \\

% \end{document}
