%%%%%%% Lecture Notes Style for Concrete Mathematics %%%%%%%%%%%%%
%%%%%%% Taught by Patrick White, TJHSST %%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,twosided]{article}
\input{concrete-header.tex}

%%%% add package inclusions here, if any
\usepackage{blubase}
%%%%% Define your custom functions and macros here, if any

%%%%%%%%%%%%%%%%%%%% Define the variables below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\titlestring{Partition Numbers}
\def\scribestring{Bryan Lu}
\def\datestring{Monday, April 8, 2019}


%%%%%%%%%%%%%%%%%%% Page Headers -- Do Not Change %%%%%%%%%%%%%%%%%%%%%%%%%%%
\lhead{\titlestring}
\rhead{Page \thepage}
\cfoot{Concrete Math -- White -- TJHSST}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

%%%%%%%%%%%%%%%%% Begin the document %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{plain}  %% no headers on this page

%%%% Do not change these lines %%%%%
\noindent
{\LARGE \textbf{\titlestring}}\\\\
%
{\Large Scribe: \scribestring}\\ \\
{\textbf{Date}: \datestring}


%%%%%%%%%%%%%%%%%%%%%%%%%% YOUR CONTENT GOES HERE %%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent

\section{Partition Numbers}
The last combinatorial numbers that we have yet to look at in our twelve-fold table are the partition numbers. We define the \textbf{partition number} $p(n)$ as the number of ways to write $n$ as a sum of positive integers. For example, $p(5) = 7$ by enumeration: 
\begin{align*}
	5 &= 1 + 1 + 1 + 1 + 1 = 1 + 1 + 1 + 2 = 1 + 2 + 2 = 1 + 1 + 3 \\
	&= 1 + 4 = 2 + 3 
\end{align*} 


Note the distinction between partitions and compositions of numbers - when doing compositions, we are counting the partitions $3, 1, 1$, $1, 3, 1$, and $1, 1, 3$ as independent ways to make compositions of $5$, while these are all the same partition of $5$.  

The explicit formula for $p(n)$ is given by 
\[
	p(n) = \frac{1}{\pi\sqrt{2}} \sum_{k=1}^\infty \sqrt{k} A_k(n) \dv{}{n}\left( \frac{1}{\sqrt{n - \frac{1}{24}}} \sinh \left[\frac{\pi}{k} \sqrt{\frac{2}{3} \left(n - \frac{1}{24}\right)} \right]\right)
\]
where $A_k$ is given by
\[
	A_k = \sum_{\gcd(m, k) = 1, m < k}e^{\pi i (S(m, k)- 2nm/k}
\]	
derived by Ramanujan.

We can still state some results regarding partition numbers by comparing this to compositions. First, let $p_k(n)$ be the number of ways to partition $n$ into $k$ parts. Notice that the number of solutions to 
\[
	x_1 + x_2 + \ldots x_k = n
\]
is given by $\tsnb{k}{n-k}$, using the multiset notation, or also $\binom{n-1}{k-1}$. Each composition contributes at most $k!$ compositions, so the number of partitions on $k$ elements, $p_k(n)$, multiplied by the number of ways to arrange $k$ distinct elements is: 
\[
	k! p_n(k) \geq \binom{n-1}{k-1}
\]
so we have that we can bound $p_k(n)$ from below by 
\[
	p_k(n) \geq \frac{1}{k!}\binom{n-1}{k-1}
\]

\section{Ferrers Graphs/Young Tableaux}
We can write partitions of positive numbers in Ferrers graphs or Young tableaux - for example, here is a partition of $7$ into 4, 2, and 1:
\begin{center}
\begin{tabular}{c c c c}
$\circ$ & $\circ$ & $\circ$ & $\circ$ \\
$\circ$ & $\circ$ &  &  \\
$\circ$ &  & & \\

\end{tabular}
\end{center}

The transpose of each of these tableaux is still a partition, $\lambda^T$. Here is the transpose of the tableaux we saw above: 
\begin{center}
\begin{tabular}{c c c}
$\circ$ & $\circ$ & $\circ$ \\
$\circ$ & $\circ$ & \\
$\circ$ &  &\\
$\circ$ &  &  \\
\end{tabular}
\end{center}

If a partition $\lambda$ is \textit{self-conjugate}, then $\lambda^T = \lambda$. Here's a theorem that we've kind of seen before: 
\begin{theorem}
The number of self-congruent partitions $\lambda$ of $n$ is the number of partitions of $n$ into \textit{distinct} odd-sized parts.
\end{theorem} 
\begin{proof}
Note that if we draw the Young tableaux for a self-conjugate partition, we can construct a partition of $n$ into distinct odd numbers by taking the "outer" row/column (ie. the left most and the topmost row) and repeating on the remaining tableaux. If we repeat this process, this gives us a partition of $n$ into odd numbers. 

To go backwards, we can construct a self-conjugate partition by constructing the Young tableaux from the inside out, starting from the smallest odd number. 

\end{proof}

Another theorem of this nature: 
\begin{theorem}
The number of partitions of $n$ into even parts is the same as the number of partitions of $n$ into even multiplicity. 
\end{theorem}
\begin{proof}
Draw out a Young tableaux representing a partition of even multiplicity. Note that the columns are even-height, so the transpose of this partition will have even-length rows.  This is easily reversible, so we have a bijection. 

\end{proof}

\begin{theorem}
\[
	p_4(n) = p_4(3n)
\]
\end{theorem}
\begin{proof}
Consider a Young tableaux in a $4 \times n$ rectangle. The area taken up by the tableaux is $n$, and the area of the rectangle is $4n$, so the remaining area, when flipped upside down, is another Young tableaux for $3n$. This completes the proof. 

\end{proof}

In general, for integers $k, n$: 
\[
	p_k(n) = p_k((k-1)n)
\]

Finally, a recursive formula: 
\[
	p_k(n) = \sum_{i=1}^k p_i(n-k)
\]

\begin{proof}
Consider chopping off the left-hand column of a Ferrers graph of $n$ that has $k$ rows, and then taking the transpose of the rows gives the recurrence for partitions of $n$ over all possibilities. 

\end{proof}














\end{document}
