%%%%%%% Lecture Notes Style for Concrete Mathematics %%%%%%%%%%%%%
%%%%%%% Taught by Patrick White, TJHSST %%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[11pt,twosided]{article}
\input{concrete-header.tex}

%%%% add package inclusions here, if any

%%%%% Define your custom functions and macros here, if any

%%%%%%%%%%%%%%%%%%%% Define the variables below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\titlestring{The Lecture Title}
\def\scribestring{Your Name}
\def\datestring{Day, Mon, Date Year}


%%%%%%%%%%%%%%%%%%% Page Headers -- Do Not Change %%%%%%%%%%%%%%%%%%%%%%%%%%%
\lhead{\titlestring}
\rhead{Page \thepage}
\cfoot{Concrete Math -- White -- TJHSST}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

%%%%%%%%%%%%%%%%% Begin the document %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\thispagestyle{plain}  %% no headers on this page

%%%% Do not change these lines %%%%%
\noindent
{\LARGE \textbf{\titlestring}}\\\\
%
{\Large Scribe: \scribestring}\\ \\
{\textbf{Date}: \datestring}


%%%%%%%%%%%%%%%%%%%%%%%%%% YOUR CONTENT GOES HERE %%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent

\section{Permutation Groups and Burnside's Lemma}
\textbf{Definition.} Let $S$ be a set and $G$ be a group of permutations $\pi$, acting on elements on $S$. Then $G$ is a \textbf{permutation group}. 

What exactly do we mean by a permutation? A permutation refers essentially to the following mapping of elements: 
\[
\pi_1 = 
\begin{pmatrix}
1 & 2 & 3 & 4 & 5 \\
3 & 2 & 1 & 5 & 4 \\
\end{pmatrix}
\]

Notice that $\pi_1(1) = 3$ and $\pi_1(4) = 5$, etc. 
We also showed in Unit 1 that every permutation can be written as a product of cycles. $\pi_1$ can thus be decomposed into cycles and written in the following way: 
\[
	\pi_1 = (13) (2) (45)
\]
This is unique up to rotations and reordering of the cycles - so the following is the same permutation: 
\[
	\pi_1 = (2) (1 3)(54)
\]
Permutations can be composed - notice that if we apply $\pi_1$ to itself we have
\[
	\pi_1 \pi_1 = \begin{pmatrix}
1 & 2 & 3 & 4 & 5 \\
1 & 2 & 3 & 4 & 5 \\
\end{pmatrix} = e
\]	
so $\pi_1 = \pi_1^{-1}$. Thus, $G = \{ e, \pi_1\}$ is a permutation group (if we check all four necessary properties). 

Another example of a permutation group - the group of rotations of a square. First, define the \textbf{symmetric group} $S_n$ as all of the $n!$ permutations of $n$ elements, so $S_4$ is essentially all permutations of the elements $A, B, C, D$. 

The group of rotations of a square can be pretty clearly to be every $90^\circ$ rotation of the square (which we will call the set $\{ e, \pi_1, \pi_2, \pi_3 \}$ where $e = \pi_0$ is the identity and a rotation of $90i$ degrees clockwise is $\pi_i$.

What about the group of all rotations of a square? In addition to $e, r_{90}, r_{180}, r_{270}$ (or $ \pi_1, \pi_2, \pi_3$, whatever), we have the reflections $H, V, L, R$, where $H$ is a reflection about a horizontal axis, $V$ is a reflection about a vertical axis, $L$ is a reflection about a main diagonal emanating from the top-left corner, and similarly for $R$. Below is the Cauchy table for the group: 
\begin{center}
\begin{tabular}{c|c c c c c c c c}
    & $e$ & $r_{90}$ & $r_{180}$ & $r_{270}$ & $V$ & $H$ & $L$ & $R$  \\ \hline 
$e$ & $e$ & $r_{90}$ & $r_{180}$ & $r_{270}$ & $V$ & $H$ & $L$ & $R$ \\ 

\end{tabular}
\end{center}

This group has a name - called $D_4$, or the dihedral group on four elements. It turns out that $|D_4| = 8$, and it turns out that $|D_n| = 2n$. 

What is one $\pi \in S_4$ that is not in $D_4$? 

Notice that we can write all the elements of $D_4$ as products of cycles: 



















\end{document}
