\setcounter{section}{-1}
\section{Introduction}
This is an archive of student notes from the Concrete Mathematics class 
from Spring 2019 at TJHSST, taught by Dr. Patrick White. 
Concrete Mathematics loosely takes from the 
textbook of the same name written by Donald Knuth, and teaches basic
combinatorial techniques and constructions with the aid of calculus.
Here, ``concrete'' is meant as a portmanteau of ``discrete'' and
``continuous,'' since we use both techniques to develop the theory (but obviously
with a focus on solving discrete problems). 
The year I took the class Dr. White 
also covered some group theory and its application in combinatorics, which may
or may not be the norm for this class. 

This class doesn't really exist anymore, but here's hoping that this semi-static
reference can serve as a historical marker as to what this class was like. It 
would be nice to see current students lobbying to bring it back, given that there is
a staff member at TJ willing to teach this course again. Personally, this class 
was the class I credit for inspiring my current interest in combinatorics.
Yes, enumerative combinatorics can seem like a grab-bag of tricks and little facts 
to know, and people from the math team will have a much easier time with the lines 
of thinking employed in this class. However, I think there is a rich theory here
that is never really explored in competitive math or in the standard calculus
pipeline in school, and this course offers a unique
opportunity to get a glimpse into what combinatorics is like at an undergraduate
level. As a bonus for the future CS majors, you'll have to take a discrete math class
anyway, and it's likely more fun here in this setting than at wherever you end up. 
Regardless, this class was a great experience and I hope that this class returns someday. 

\subsection{Editing Notes}
These notes were written by the class, with a rotating scribing system to record
what happened in lectures every week, and were either transcribed or copied by me
from the original saved tex/pdf sources. The previous page gives credit to the
people who scribed for the class on which days. Here are my notes on editing: 
\begin{itemize}
    \item I tried to be faithful as possible to the source -- either by essentially including 
    the tex source as input or by transcribing the pdf as best as I could (while also potentially
    fixing some errors along the way).
    \item If two sets of notes existed on a day, contributions from both authors are 
    kept in the repository, but the one in this main document is the one that I found to be
    more coherent/more thorough than the other. All authors are still credited for every day, 
    regardless if their words are the ones that are in the main document. 
    \item There are some jumps in the dates listed on the previous page 
    -- Mondays sometimes don't exist? I don't really remember why
    to be honest... maybe no lectures happened those days because of quizzes or some other 
    reason? Anyways, the dates are fairly arbitrary and can be safely ignored if you wish. 
    \item Some contributors' graduation years/personal information may be wrong -- I tried to 
    put people's current preferred names and graduation years as I remembered them, but I'm 
    not sure I got everyone right. 
\end{itemize}
Errors, corrections, etc. are welcome and you can email me or submit a pull request if you like.